Canonical rings of $\bq$-divisors on Minimal Rational Surfaces

### 2015-07-28 surfaces-write-up/surface-write-up.tex
responses to peter's hirzebruch comments and more comments that are confusing me

\begin{defn}\label{def:ell_i}
Let $D = \sum_{i=1}^n \alpha_i D_i$ where $\alpha_i = \frac{c_i}{k_i}$. \todo{Peter: better notation for fraction? Aaron: This seems fine to me. Are you saying you want a better notation, or asking if this is a better notation?}  Then notate
\[\ell_i = \lcm_{j\ne i}(k_i).\]
\end{defn}

@@

Let $\omega\in \Omega'$ correspond to the monomial $y_i \prod_{j=1}^n
{y_j}^{c_j}$.  Then we can write $\omega$ as a sum of $r_j e_j$'s for $j\in
\Te$ and $r_{j,k}e_{j,k}$ for $j\in \Tp, k\in \Tm$.  
For convenience, let $e_j$ (when it exists) be in degree $d_j$ with $j^{th}$
component $-\alpha_j d_j + \kappa_j$ for some $\kappa_j \in \bq$ and let $e_{j,k}$ (when it exists) lie in
degrees $d_{j,k}$ with $j^{th}$ component $-\alpha_j d_{j,k} + \kappa_{j,k}'$ and $k^{th}$
component $-\alpha_j d_{j,k} + \kappa_{j,k}''$ for $\kappa_{j,k}', \kappa_{j,k}'' \in \bq$
By a careful computation, we can write

\todo{Aaron: What is $\sigma$ here? Is it $\rho$? \\
Peter: Should have been $\omega$...}
%if $i\in \Te$, $i\in \Tp$, and $i\in \Tm$, then
\begin{equation}\label{eqn:sigma-in-terms-of-lambda/E}
	\omega = \lambda + E
\end{equation}
for some $\lambda \in \Sigma$ and $E$ of the form
\begin{equation}\label{eqn:hirz-E-translation}
	E = \begin{cases}
	\frac{1}{\kappa_i} e_i  	&  \mbox{ if }i\in \Te\\
	\sum_{j\in \Tm} r_{i,j} e_{i,j}: r_{i,j}\in \bq, \sum_{j\in \Tm} r_{i,j}\kappa_{i,j}' = 1 & \mbox{ if } i\in \Tp\\
	\sum_{j\in \Tp} r_{j,i} e_{j,i}: r_{j,i} \in \bq, \sum_{j\in \Tm} r_{j,i} \kappa_{i,j}'' = 1 	& \mbox{ if } i\in \Tm\\
	\end{cases}.
\end{equation}

\noindent
From this, we can further deduce that for such an $E$,
\[
	\deg(E) \le \begin{cases}
	\ell_i \gcd(\bida_i, \bidb_i)	&\mbox{ if } i \in \Te \\
	\max_{j \in \Tm} \bigl(\ell_{i, j} (\bida_i \bidb_j - \bida_j \bidb_i)\bigr)
	&\mbox{ if } i \in \Tp \\
	\max_{j \in \Tp} \bigl(\ell_{j, i} (\bida_j \bidb_ i - \bida_i \bidb_j) \bigr)
	&\mbox{ if } i \in \Tm \end{cases}.
\]
Let $M$ be the set of monomials terms of $\beta_i$.  Then for $\mu = \prod_{j=1}^n
{f_i}^{c_i}\in M$, we
\todo{Aaron: What is $M$? \\
Peter: I don't quite follow this question... Is your objection that ``the set of monomial terms of $\beta_i$'' is too imprecise?}
\todo{Aaron: Again, why do we keep defining lattce points in
	terms of elements in the canonical ring? Why not just
write out what the lattice points are in coordinates? \\ 
Peter: Fixed?}
can further compute that the lattice point, $\omega_\mu = (c_1 ,\ldots c_n)$, corresponding to $\mu
$ is of the form
\[
	\omega_\mu = \lambda + E_\mu
\]
where $\lambda \in \Sigma$ is the same as in Equation \ref{eqn:sigma-in-terms-of-lambda/E} and $E_\mu$ is defined in a similar fashion as $E$ in Equation \ref{eqn:hirz-E-translation}:
\begin{equation}\label{eqn:hirz-E_mu-translation}
E_\mu = \left(\sum_{\substack{i \in \{1, 2, 3, 4\} \cap \Te}}q_i e_i\right) + \left(\sum_{\substack{i \in \{1, 2, 3, 4\} \cap \Tp \\ j\in \Tm}} q_{i,j} e_{i,j}\right) + \left(\sum_{\substack{i \in \{1, 2, 3, 4\} \cap \Tm \\ j\in \Tp}} q_{j,i} e_{i,j}\right)
\end{equation}
for some $q_i$'s, $q_{i,j}$'s and $q_{j,i}$'s (for appropriate $i$'s and $j$'s) in $\bq_{\ge 0}$.  
\todo{Aaron: if $\sigma_\mu$ is an element, how can it be written as
	the sum of an element and a set? Same question for $\sigma$ above.
	Also, I do not know what $E_\mu$ is. You really have to define it,
since there's no way to check the validity of this proof without it.  \\
Peter: $E$ and $E_\mu$ are elements, not sets...}
Note that $E_\mu$ lies in the same degree as $E$ for all $\mu$.

@@

\todo{add something about ghost points?}
\todo{Aaron: I know exactly what is going on, but please spell out the proof in 2 lines for completeness.\\
Peter: Better?}

>>>>

\begin{defn}\label{def:ell_i}
Let $D = \sum_{i=1}^n \alpha_i D_i$ where $\alpha_i = \frac{c_i}{k_i}$.
Then notate
\[\ell_i = \lcm_{j\ne i}(k_i).\]
\end{defn}

@@

Let $\omega\in \Omega'$ correspond to the monomial $y_i \prod_{j=1}^n
{y_j}^{c_j}$.  Then we can write $\omega$ as a sum of $r_j e_j$'s for $j\in
\Te$ and $r_{j,k}e_{j,k}$ for $j\in \Tp, k\in \Tm$.  
For convenience, let $e_j$ (when it exists) be in degree $d_j$ with $j^{th}$
component $-\alpha_j d_j + \kappa_j$ for some $\kappa_j \in \bq$ and let $e_{j,k}$ (when it exists) lie in
degrees $d_{j,k}$ with $j^{th}$ component $-\alpha_j d_{j,k} + \kappa_{j,k}'$ and $k^{th}$
component $-\alpha_j d_{j,k} + \kappa_{j,k}''$ for $\kappa_{j,k}', \kappa_{j,k}'' \in \bq$
By a careful computation, we can find $\lambda \in \Sigma$ 
\todo{Aaron: Is it at all possible to write down what $\lambda$ and the $r_{i,j}$ are?
I suspect we should be able to write them in a couple lines, which would
definitely clarify how we're getting the bound on degree $E$. While I agree
it's fine to hide the computation of the $r_{i,j}$ it would still clarify matters
to say what the $r_{i,j}$ and $\lambda$ are.}
and $r_{i,j} \in \bq$ so that we can write
\begin{equation}\label{eqn:sigma-in-terms-of-lambda/E}
-	\sigma = \lambda + E
 \end{equation}
with $E$ of the form
\begin{equation}\label{eqn:hirz-E-translation}
	E = \begin{cases}
	\frac{1}{\kappa_i} e_i  	&  \mbox{ if }i\in \Te\\
	\sum_{j\in \Tm} r_{i,j} e_{i,j}: r_{i,j}\in \bq, \sum_{j\in \Tm} r_{i,j}\kappa_{i,j}' = 1 & \mbox{ if } i\in \Tp\\
	\sum_{j\in \Tp} r_{j,i} e_{j,i}: r_{j,i} \in \bq, \sum_{j\in \Tm} r_{j,i} \kappa_{i,j}'' = 1 	& \mbox{ if } i\in \Tm\\
	\end{cases}.
\end{equation}

\noindent
From this, we can further deduce that for such an $E$,
\[
	\deg(E) \le \begin{cases}
	\ell_i \gcd(\bida_i, \bidb_i)	&\mbox{ if } i \in \Te \\
	\max_{j \in \Tm} \bigl(\ell_{i, j} (\bida_i \bidb_j - \bida_j \bidb_i)\bigr)
	&\mbox{ if } i \in \Tp \\
	\max_{j \in \Tp} \bigl(\ell_{j, i} (\bida_j \bidb_ i - \bida_i \bidb_j) \bigr)
	&\mbox{ if } i \in \Tm \end{cases}.
\]
Let $M$ be the set of monomials terms of $\beta_i$.  Then for $\mu = \prod_{j=1}^n
{f_i}^{c_i}\in M$, we
\todo{Aaron: What is $M$? \\
Peter: I don't quite follow this question... Is your objection that ``the set of monomial terms of $\beta_i$'' is too imprecise?

Aaron: Sorry, I was trying to say that
I don't understand what ``the set of monomials in terms of beta i'' means.
Do you just mean that $M = \{\psi(\beta_1), \ldots, \psi(\beta_n)\}$?

However, I'm really baffled as to why you want to apply $\psi$ to the
$\beta_i$, since two lines later, you then apply the reverse correspondence
and want to work back in the semigroupring. Why don't you just stay in the
semigroup ring the whole time?
I think I'm misunderstanding this, but is $\omega_\mu$ just equal to
$\beta_i$ for some $i$? Also, why is $\omega_\mu$ missing a $d$ index
as its first index?
}
\todo{Aaron: Again, why do we keep defining lattce points in
	terms of elements in the canonical ring? Why not just
write out what the lattice points are in coordinates? \\ 
Peter: Fixed?

Aaron: I like that $E$ is defined in terms of $e_i$.
I'm still confused about $\omega$ and $\mu$ though. see my above comment.
}
can further compute that there exist $q_i, q_{i,j} \in \bq_{\geq 0}$ and $\lambda \in
\Sigma$ so that the lattice point, $\omega_\mu = (c_1 ,\ldots c_n)$,
corresponding to $\mu
$ is of the form
\[
	\omega_\mu = \lambda + E_\mu
\]
where $\lambda$ is the same as in Equation \ref{eqn:sigma-in-terms-of-lambda/E} and $E_\mu$ is defined in a similar fashion as $E$ in Equation \ref{eqn:hirz-E-translation} by
\begin{equation}\label{eqn:hirz-E_mu-translation}
E_\mu = \left(\sum_{\substack{i \in \{1, 2, 3, 4\} \cap \Te}}q_i e_i\right) +
\left(\sum_{\substack{i \in \{1, 2, 3, 4\} \cap \Tp \\ j\in \Tm}} q_{i,j}
e_{i,j}\right) + \left(\sum_{\substack{i \in \{1, 2, 3, 4\} \cap \Tm \\ j\in
\Tp}} q_{j,i} e_{i,j}\right).
\end{equation}
\todo{Aaron: if $\sigma_\mu$ is an element, how can it be written as
	the sum of an element and a set? Same question for $\sigma$ above.
	Also, I do not know what $E_\mu$ is. You really have to define it,
since there's no way to check the validity of this proof without it.  \\
Peter: $E$ and $E_\mu$ are elements, not sets...

Aaron: My mistake. Thanks for writing it out though. However,
since $E, E_\mu$ are so similar, why not defining them so that they look similar?
I.e., why not define them either both in three cases as $E$
or both as sums as $E_\mu$. It's confusing that they are the same, but not
written in a parallel fashion.
}
Note that $E_\mu$ lies in the same degree as $E$ for all $\mu$.

@@

\todo{add something about ghost points?}
